---
title: "WASH Benefits Bangladesh Analysis of Stress Biomarkers and Growth Outcomes"
author: "Andrew Mertens"
date: "05/14/2024"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

source(here::here("0-config.R"))

library(tidyverse)
library(knitr)
#library(boxr)
library(here)
library(scales)
#box_auth()

theme_ki<-function(){
  theme_bw() %+replace%
    theme(
      strip.background = element_blank(),
      legend.position="none",
      plot.title = element_text(size = 16, face = "bold"),
      strip.text = element_text(size=14),
      axis.title = element_text(size=12),
      axis.text.y = element_text(size=10),
      axis.text.x = element_text(size=10, angle = 0, hjust = 0.5, vjust=.1)
    )
}

theme_set(theme_ki())


clean_H1 <- function(res){
  res$Y <- gsub("_t2_t3","",res$Y)
  res$Y <- gsub("delta","Change in",res$Y)
  res$Y <- gsub("_"," ",res$Y)
  
  res$Y <- gsub("laz","LAZ",res$Y)
  res$Y <- gsub("waz","WAZ",res$Y)
  res$Y <- gsub("whz","WLZ",res$Y)
  res$Y <- gsub("hcz","HCZ",res$Y)
  
  res$Y <- gsub("t2","-Year 1",res$Y)
  res$Y <- gsub("t3","-Year 2",res$Y)
  
  res$Y <- gsub("len","Length",res$Y)
  res$Y <- gsub("wei","Weight",res$Y)
  res$Y <- gsub("hc","Head Circumference",res$Y)
  
  unique(res$Y)
  res$Y <- factor(res$Y, levels = rev(c("LAZ -Year 1","WAZ -Year 1","WLZ -Year 1","HCZ -Year 1",               
                                        "LAZ -Year 2",   "WAZ -Year 2","WLZ -Year 2","HCZ -Year 2",                
                                        "Change in LAZ","Change in WAZ","Change in WLZ","Change in HCZ",
                                        "Length velocity","Weight velocity","Head Circumference velocity")))
  return(res)
}

tableau10 <- c("#1F77B4","#FF7F0E","#2CA02C","#D62728",
               "#9467BD","#8C564B","#E377C2","#7F7F7F",
               "#BCBD22","#17BECF")

dropboxDir <- "C:/Users/andre/Dropbox/WASHB-EE-analysis/WBB-EE-analysis/"


#heatmaps
load(paste0(here::here(),"/figures/supplementary/heatmaps.Rdata"))


H1_res <- readRDS(paste0(here::here(),"/results/adjusted/H1_adj_res.RDS"))
H2_res <- readRDS(paste0(here::here(),"/results/adjusted/H2_adj_res.RDS"))
H3_res <- readRDS(paste0(here::here(),"/results/adjusted/H3_adj_res.RDS"))
H4_res <- readRDS(paste0(here::here(),"/results/adjusted/H4_adj_res.RDS"))

H1_plot_data <- readRDS(paste0(dropboxDir,"results/stress-growth-models/figure-data/H1_adj_spline_data.RDS"))
H2_plot_data <- readRDS(paste0(dropboxDir,"results/stress-growth-models/figure-data/H2_adj_spline_data.RDS"))
H3_plot_data <- readRDS(paste0(dropboxDir,"results/stress-growth-models/figure-data/H3_adj_spline_data.RDS"))
H4_plot_data <- readRDS(paste0(dropboxDir,"results/stress-growth-models/figure-data/H4_adj_spline_data.RDS"))
boxplot <- readRDS(paste0(dropboxDir,"results/stress-growth-models/figure-objects/boxplot_objects.RDS"))




dfull <- readRDS("C:/Users/andre/Documents//EE/eed-substudy-data/bangladesh-cleaned-master-data.RDS")


d <- dfull %>% select(childid, t2_f2_23d, t2_f2_23d_raw, laz_t1,laz_t2,laz_t3)
d1 <- d %>% select(childid, t2_f2_23d, t2_f2_23d_raw, laz_t1) %>% rename(laz=laz_t1) %>% mutate(time="3 months")
d2 <- d %>% select(childid, t2_f2_23d, t2_f2_23d_raw, laz_t2) %>% rename(laz=laz_t2) %>% mutate(time="14 months")
d3 <- d %>% select(childid, t2_f2_23d, t2_f2_23d_raw, laz_t3) %>% rename(laz=laz_t3) %>% mutate(time="28 months")

d <- bind_rows(d1,d2,d3) %>% mutate(time=factor(time, levels=c("3 months","14 months","28 months"))) %>%
  filter(!is.na(t2_f2_23d))

#df <- d %>% filter(time!="3 months") %>% group_by(childid) %>% mutate(N=n()) %>% filter(N==2)
df <- d %>% #filter(time!="3 months") %>% 
  group_by(childid) %>% mutate(N=n()) %>% filter(N==3)
df <- df %>% ungroup() %>% mutate(f2_cat=ntile(t2_f2_23d,4))
df2 <- df %>% group_by(childid) %>% mutate(delta_laz=round(last(laz)-first(laz),2))




```



## Primary findings - HAZ


```{r}



H1_res$Xtime <- "t2"
H1_res <- H1_res %>% mutate(
  Ytime = case_when(
    grepl("_t2_t3",H1_res$Y) ~ "delta",
    grepl("_t2",H1_res$Y) ~ "t2",
    grepl("_t3",H1_res$Y) ~ "t3"),
Ytype = case_when(
  grepl("laz",H1_res$Y) ~ "LAZ",
  grepl("waz",H1_res$Y) ~ "WAZ",
  grepl("whz",H1_res$Y) ~ "WHZ",
  grepl("hcz",H1_res$Y) ~ "HCZ"))
H1_res$sig.cat <- ifelse(H1_res$BH.Pval < 0.05, "*", "")


plotdf <- H1_res %>% filter(Ytype=="LAZ") %>% droplevels()  %>%
  mutate(Y=factor(Y, levels=c("laz_t2","laz_t3","delta_laz_t2_t3")))

p <- ggplot(data = plotdf, (aes(x=X, y=point.diff#,
                                #group=sample_cat, color=sample_cat
                                ))) +
  geom_point(size=3, position = position_dodge(0.5), alpha=0.75) +
  geom_errorbar(aes(ymin=lb.diff, ymax=ub.diff), position = position_dodge(0.5),
                width = 0.3, size = 1) +
  facet_wrap(~Y,  scales="free_x") +
  geom_text(aes(label=sig.cat), color="black", position = position_dodge(0.5), vjust = -0.1) +
  coord_flip() +
  # scale_color_manual(breaks = legend_labels,
  #                    values = colours, drop = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed") +
  #labs(color="Sample type") +
  xlab("Stress biomarker") + ylab("Mean difference: 25th vs. 75th percentile") +
  theme_ki() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "bottom",
        strip.placement = "outside",
        strip.text.x = element_text(size=10, face = "bold"),
        strip.text.y = element_text(size=10, angle = 270, face = "bold"),          plot.title = element_text(hjust = 0.5, face = "plain", size=9),
        panel.spacing = unit(0, "lines"))
plot(p)

```



## Exposure-response: Concurrent f2 8ip and LAZ

```{r}
H1_plot_data_sub <- H1_plot_data %>% filter(Xvar=="t2_f2_8ip", Yvar=="laz_t2")
gam_diff <- H1_res %>% filter(X=="t2_f2_8ip", Y=="laz_t2")
p <- ggplot(H1_plot_data_sub) +
  geom_ribbon(aes(x = X, ymin = lwrS, ymax = uprS),
      alpha = 0.5) +
  geom_vline(aes(xintercept=q1), data=gam_diff) +
  geom_vline(aes(xintercept=q3), data=gam_diff) +
  geom_path(aes(x = X, y = fit), color = "black") +
  xlab("f2_8ip") + ylab("LAZ") +
  #ggtitle("Concurrent association")+
  ggtitle(paste0(round(gam_diff$point.diff, 2), " (",
                 round(gam_diff$lb.diff, 2), ", ", round(gam_diff$ub.diff,
                                                         2), ")"))
p


```

## Exposure-response: Concurrent f2 8ip and LAZ

```{r}
p <- ggplot(H1_plot_data_sub) +
  geom_point(aes(x = X, y = Y), alpha=0.2) +
  geom_ribbon(aes(x = X, ymin = lwrS, ymax = uprS), alpha = 0.5) +
  geom_path(aes(x = X, y = fit), color = "black", size=1.5) +
  geom_vline(aes(xintercept=q1), data=gam_diff) +
  geom_vline(aes(xintercept=q3), data=gam_diff) +
  xlab("") + ylab("") +
  ggtitle("")
p
```


## Exposure-response: Concurrent f2 23d and LAZ

```{r}

H1_plot_data_sub <- H1_plot_data %>% filter(Xvar=="t2_f2_23d", Yvar=="laz_t2")
gam_diff <- H1_res %>% filter(X=="t2_f2_23d", Y=="laz_t2")
p <- ggplot(H1_plot_data_sub) +
  geom_ribbon(aes(x = X, ymin = lwrS, ymax = uprS),
      alpha = 0.5) +
  geom_vline(aes(xintercept=q1), data=gam_diff) +
  geom_vline(aes(xintercept=q3), data=gam_diff) +
  geom_path(aes(x = X, y = fit), color = "black") +
  xlab("f2_23d") + ylab("LAZ") +
  #ggtitle("Concurrent association")+
  ggtitle(paste0(round(gam_diff$point.diff, 2), " (",
                 round(gam_diff$lb.diff, 2), ", ", round(gam_diff$ub.diff,
                                                         2), ")"))
p

```


## Exposure-response: Concurrent f2 VI and LAZ

```{r}
H1_plot_data_sub <- H1_plot_data %>% filter(Xvar=="t2_f2_VI", Yvar=="laz_t2")
gam_diff <- H1_res %>% filter(X=="t2_f2_VI", Y=="laz_t2")
p <- ggplot(H1_plot_data_sub) +
  geom_ribbon(aes(x = X, ymin = lwrS, ymax = uprS),
      alpha = 0.5) +
  geom_vline(aes(xintercept=q1), data=gam_diff) +
  geom_vline(aes(xintercept=q3), data=gam_diff) +
  geom_path(aes(x = X, y = fit), color = "black") +
  xlab("f2 VI") + ylab("LAZ") +
  #ggtitle("Concurrent association")+
  ggtitle(paste0(round(gam_diff$point.diff, 2), " (",
                 round(gam_diff$lb.diff, 2), ", ", round(gam_diff$ub.diff,
                                                         2), ")"))
p

```


## Exposure-response: Concurrent f2 iso pca and LAZ

```{r}
H1_plot_data_sub <- H1_plot_data %>% filter(Xvar=="t2_iso_pca", Yvar=="laz_t2")
gam_diff <- H1_res %>% filter(X=="t2_iso_pca", Y=="laz_t2")
p <- ggplot(H1_plot_data_sub) +
  geom_ribbon(aes(x = X, ymin = lwrS, ymax = uprS),
      alpha = 0.5) +
  geom_vline(aes(xintercept=q1), data=gam_diff) +
  geom_vline(aes(xintercept=q3), data=gam_diff) +
  geom_path(aes(x = X, y = fit), color = "black") +
  xlab("f2 iso pca") + ylab("LAZ") +
  #ggtitle("Concurrent association")+
  ggtitle(paste0(round(gam_diff$point.diff, 2), " (",
                 round(gam_diff$lb.diff, 2), ", ", round(gam_diff$ub.diff,
                                                         2), ")"))
p

```



## Primary findings - HCZ


```{r}


plotdf <- H1_res %>% filter(Ytype=="HCZ") %>% droplevels() %>%
  mutate(Y=factor(Y, levels=c("hcz_t2","hcz_t3","delta_hcz_t2_t3")))
p <- ggplot(data = plotdf, (aes(x=X, y=point.diff#,
                                #group=sample_cat, color=sample_cat
                                ))) +
  geom_point(size=3, position = position_dodge(0.5), alpha=0.75) +
  geom_errorbar(aes(ymin=lb.diff, ymax=ub.diff), position = position_dodge(0.5),
                width = 0.3, size = 1) +
  facet_wrap(~Y,  scales="free_x") +
  geom_text(aes(label=sig.cat), color="black", position = position_dodge(0.5), vjust = -0.1) +
  coord_flip() +
  # scale_color_manual(breaks = legend_labels,
  #                    values = colours, drop = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed") +
  #labs(color="Sample type") +
  xlab("Stress biomarker") + ylab("Mean difference: 25th vs. 75th percentile") +
  theme_ki() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "bottom",
        strip.placement = "outside",
        strip.text.x = element_text(size=10, face = "bold"),
        strip.text.y = element_text(size=10, angle = 270, face = "bold"),          plot.title = element_text(hjust = 0.5, face = "plain", size=9),
        panel.spacing = unit(0, "lines"))
plot(p)

```

## Primary findings - WAZ


```{r}



plotdf <- H1_res %>% filter(Ytype=="WAZ") %>% droplevels() %>%
  mutate(Y=factor(Y, levels=c("waz_t2","waz_t3","delta_waz_t2_t3")))
p <- ggplot(data = plotdf, (aes(x=X, y=point.diff#,
                                #group=sample_cat, color=sample_cat
                                ))) +
  geom_point(size=3, position = position_dodge(0.5), alpha=0.75) +
  geom_errorbar(aes(ymin=lb.diff, ymax=ub.diff), position = position_dodge(0.5),
                width = 0.3, size = 1) +
  facet_wrap(~Y,  scales="free_x") +
  geom_text(aes(label=sig.cat), color="black", position = position_dodge(0.5), vjust = -0.1) +
  coord_flip() +
  # scale_color_manual(breaks = legend_labels,
  #                    values = colours, drop = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed") +
  #labs(color="Sample type") +
  xlab("Stress biomarker") + ylab("Mean difference: 25th vs. 75th percentile") +
  theme_ki() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "bottom",
        strip.placement = "outside",
        strip.text.x = element_text(size=10, face = "bold"),
        strip.text.y = element_text(size=10, angle = 270, face = "bold"),          plot.title = element_text(hjust = 0.5, face = "plain", size=9),
        panel.spacing = unit(0, "lines"))
plot(p)

```

## Primary findings - WHZ


```{r}


plotdf <- H1_res %>% filter(Ytype=="WHZ") %>% droplevels() %>%
  mutate(Y=factor(Y, levels=c("whz_t2","whz_t3","delta_whz_t2_t3")))
p <- ggplot(data = plotdf, (aes(x=X, y=point.diff#,
                                #group=sample_cat, color=sample_cat
                                ))) +
  geom_point(size=3, position = position_dodge(0.5), alpha=0.75) +
  geom_errorbar(aes(ymin=lb.diff, ymax=ub.diff), position = position_dodge(0.5),
                width = 0.3, size = 1) +
  facet_wrap(~Y,  scales="free_x") +
  geom_text(aes(label=sig.cat), color="black", position = position_dodge(0.5), vjust = -0.1) +
  coord_flip() +
  # scale_color_manual(breaks = legend_labels,
  #                    values = colours, drop = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed") +
  #labs(color="Sample type") +
  xlab("Stress biomarker") + ylab("Mean difference: 25th vs. 75th percentile") +
  theme_ki() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "bottom",
        strip.placement = "outside",
        strip.text.x = element_text(size=10, face = "bold"),
        strip.text.y = element_text(size=10, angle = 270, face = "bold"),          plot.title = element_text(hjust = 0.5, face = "plain", size=9),
        panel.spacing = unit(0, "lines"))
plot(p)

```


## Primary findings: synthesized in a heatmap

```{r}

print(hm1)

```

## Sensitivity analysis: unadjusted estimates

```{r}

print(hm1_unadj)

```

## Sensitivity analysis: linear regression results

```{r}

print(hm1_glm)

```


## Descriptive results: boxplots of isoprotanes by growth category



```{r}

print(boxplot)

```

## Primary findings: H2

No associations significant after P-value correction with other growth outcomes

```{r}


H2_res$Xtime <- "t2"
H2_res <- H2_res %>% mutate(
  Ytime = case_when(
    grepl("_t2_t3",H2_res$Y) ~ "delta",
    grepl("_t2",H2_res$Y) ~ "t2",
    grepl("_t3",H2_res$Y) ~ "t3"),
Ytype = case_when(
  grepl("laz",H2_res$Y) ~ "LAZ",
  grepl("waz",H2_res$Y) ~ "WAZ",
  grepl("whz",H2_res$Y) ~ "WHZ",
  grepl("hcz",H2_res$Y) ~ "HCZ"))
H2_res$sig.cat <- ifelse(H2_res$BH.Pval < 0.05, "*", "")

plotdf <- H2_res %>% filter(Ytype=="LAZ") %>% droplevels() %>%
  mutate(Y=factor(Y, levels=c("laz_t2","laz_t3","delta_laz_t2_t3")))
p <- ggplot(data = plotdf, (aes(x=X, y=point.diff#,
                                #group=sample_cat, color=sample_cat
                                ))) +
  geom_point(size=3, position = position_dodge(0.5), alpha=0.75) +
  geom_errorbar(aes(ymin=lb.diff, ymax=ub.diff), position = position_dodge(0.5),
                width = 0.3, size = 1) +
  facet_wrap(~Y,  scales="free_x") +
  geom_text(aes(label=sig.cat), color="black", position = position_dodge(0.5), vjust = -0.1) +
  coord_flip() +
  # scale_color_manual(breaks = legend_labels,
  #                    values = colours, drop = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed") +
  #labs(color="Sample type") +
  xlab("Stress biomarker") + ylab("Mean difference: 25th vs. 75th percentile") +
  theme_ki() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "bottom",
        strip.placement = "outside",
        strip.text.x = element_text(size=10, face = "bold"),
        strip.text.y = element_text(size=10, angle = 270, face = "bold"),          plot.title = element_text(hjust = 0.5, face = "plain", size=9),
        panel.spacing = unit(0, "lines"))
plot(p)

```


## Primary findings: H3

No associations significant after P-value correction with other growth outcomes

```{r}


H3_res$Xtime <- "t2"
H3_res <- H3_res %>% mutate(
  Ytime = case_when(
    grepl("_t2_t3",H3_res$Y) ~ "delta",
    grepl("_t2",H3_res$Y) ~ "t2",
    grepl("_t3",H3_res$Y) ~ "t3"),
Ytype = case_when(
  grepl("laz",H3_res$Y) ~ "LAZ",
  grepl("waz",H3_res$Y) ~ "WAZ",
  grepl("whz",H3_res$Y) ~ "WHZ",
  grepl("hcz",H3_res$Y) ~ "HCZ"))
H3_res$sig.cat <- ifelse(H3_res$BH.Pval < 0.05, "*", "")

plotdf <- H3_res %>% filter(Ytype=="LAZ") %>% droplevels() %>%
  mutate(Y=factor(Y, levels=c("laz_t2","laz_t3","delta_laz_t2_t3")))
p <- ggplot(data = plotdf, (aes(x=X, y=point.diff#,
                                #group=sample_cat, color=sample_cat
                                ))) +
  geom_point(size=3, position = position_dodge(0.5), alpha=0.75) +
  geom_errorbar(aes(ymin=lb.diff, ymax=ub.diff), position = position_dodge(0.5),
                width = 0.3, size = 1) +
  facet_wrap(~Y,  scales="free_x") +
  geom_text(aes(label=sig.cat), color="black", position = position_dodge(0.5), vjust = -0.1) +
  coord_flip() +
  # scale_color_manual(breaks = legend_labels,
  #                    values = colours, drop = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed") +
  #labs(color="Sample type") +
  xlab("Stress biomarker") + ylab("Mean difference: 25th vs. 75th percentile") +
  theme_ki() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "bottom",
        strip.placement = "outside",
        strip.text.x = element_text(size=10, face = "bold"),
        strip.text.y = element_text(size=10, angle = 270, face = "bold"),          plot.title = element_text(hjust = 0.5, face = "plain", size=9),
        panel.spacing = unit(0, "lines"))
plot(p)

```


## Primary findings: H4

No associations significant after P-value correction with other growth outcomes

```{r}


H4_res$Xtime <- "t2"
H4_res <- H4_res %>% mutate(
  Ytime = case_when(
    grepl("_t2_t3",H4_res$Y) ~ "delta",
    grepl("_t2",H4_res$Y) ~ "t2",
    grepl("_t3",H4_res$Y) ~ "t3"),
Ytype = case_when(
  grepl("laz",H4_res$Y) ~ "LAZ",
  grepl("waz",H4_res$Y) ~ "WAZ",
  grepl("whz",H4_res$Y) ~ "WHZ",
  grepl("hcz",H4_res$Y) ~ "HCZ"))
H4_res$sig.cat <- ifelse(H4_res$BH.Pval < 0.05, "*", "")

plotdf <- H4_res %>% filter(Ytype=="LAZ") %>% droplevels() %>%
  mutate(Y=factor(Y, levels=c("laz_t2","laz_t3","delta_laz_t2_t3")))
p <- ggplot(data = plotdf, (aes(x=X, y=point.diff#,
                                #group=sample_cat, color=sample_cat
                                ))) +
  geom_point(size=3, position = position_dodge(0.5), alpha=0.75) +
  geom_errorbar(aes(ymin=lb.diff, ymax=ub.diff), position = position_dodge(0.5),
                width = 0.3, size = 1) +
  facet_wrap(~Y,  scales="free_x") +
  geom_text(aes(label=sig.cat), color="black", position = position_dodge(0.5), vjust = -0.1) +
  coord_flip() +
  # scale_color_manual(breaks = legend_labels,
  #                    values = colours, drop = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed") +
  #labs(color="Sample type") +
  xlab("Stress biomarker") + ylab("Mean difference: 25th vs. 75th percentile") +
  theme_ki() +
  theme(axis.ticks.x=element_blank(),
        legend.position = "bottom",
        strip.placement = "outside",
        strip.text.x = element_text(size=10, face = "bold"),
        strip.text.y = element_text(size=10, angle = 270, face = "bold"),          plot.title = element_text(hjust = 0.5, face = "plain", size=9),
        panel.spacing = unit(0, "lines"))
plot(p)

```




